### 2018-06-18 algorithm.tex
start writing up bellman

\documentclass{article}
\usepackage{amsmath}
\begin{document}
Solving a simple Armington-endowment model.

@@

Start from a candidate labor allocation $L_{njt}^{(0)}$. Calculate expected wage share as the simple average of wage shares across the $S$ realizations. Set $L_{njt}^{(n+1)} = L_{nt}\eta_{njt}^{(n)}$, where $\eta_{njt}^{(n)}$ is the wage share in iteration $n$. This may be dampened, but in practice, it convergence in a handful of steps.
\end{document}

>>>>

\documentclass[12pt]{article}
\usepackage{amsmath}
\begin{document}
Solving a simple Armington-endowment model.

@@

Start from a candidate labor allocation $L_{njt}^{(0)}$. Calculate expected wage share as the simple average of wage shares across the $S$ realizations. Set $L_{njt}^{(n+1)} = L_{nt}\eta_{njt}^{(n)}$, where $\eta_{njt}^{(n)}$ is the wage share in iteration $n$. This may be dampened, but in practice, it convergence in a handful of steps.

\subsection{Time series}
Given the autoregressive nature of productivity, the vector of $A_{jnt}$s is a state variable. The problem is drastically simplified by the fact that no decision variable affects the state. The representative agent solves
\[
\max_{\{L_{njt}\}}
E
\sum_{t=0}^{\infty} 
	\delta^{t} 
	\left[
	\ln \frac 
		{\sum_j w_{njt}(\mathbf A_{t}, \mathbf L_{t})L_{njt}}
		{P_{nt}(\mathbf A_{t}, \mathbf L_{t})}
	-\frac \varrho 2
		\sum_{j=1}^J
		\left(
			\frac {L_{njt}} {L_{nt}}
			- \psi_j^*(\mathbf A_{t})
		\right)^2
	\right]
\]
subject to $\sum_j L_{njt}=L_{nt}$ and $w$ and $P$ being determined in equilibrium. (This includes adjustment costs)

The corresponding Bellman equation is
\begin{equation}
	V_n^N(\mathbf A_{t-1}) = 
	E V_n^D(\mathbf A_{t-1}, \mathbf A_{t})
\end{equation}
for the ``night'' period before productivity shocks have been realized. For the ``day'' period,
\begin{equation}
	V_n^D(\mathbf A_{t-1}, \mathbf A_{t}) = 
	\max_{\{L_{njt}\}}
		\ln \frac 
			{\sum_j w_{njt}(\mathbf A_{t}, \mathbf L_{t})L_{njt}}
			{P_{nt}(\mathbf A_{t}, \mathbf L_{t})}
		-\frac \varrho 2
			\sum_{j=1}^J
			\left(
				\frac {L_{njt}} {L_{nt}}
				- \psi_j^*(\mathbf A_{t-1})
			\right)^2
	+ \delta E V_n^N(\mathbf A_{t}).
\end{equation}
The day value depends on both time-$t$ and time-$t-1$ productivity because $\psi^*$ has been determined during the night period, before $A_t$ was known.

Note that choice $L_{njt}$ does not affect future state $A_t$.
\end{document}
